---
title: "HW-02 - Data Wrangling and Visualization"
author: "Your name"
date: "today's date"
output: 
  html_document: 
    theme: cerulean
---

# Part I:  College Graduation Rates
The data set `colleges.csv` contains information about 1991 graduation rates for the top 200 American universities and colleges determined by "Money" magazine. 


1.  The `colleges.csv` data set can be found in the `/data` folder.  Read in the data and add a new variable that estimates the tuition for each college in the year 2000.  The tuition inflation rate is about 8% per year, which means that the cost of college doubles every nine years.  Create another variable that calculates the tuition *increase* from 1991 to 2000 for each college.  Create a visualization for the distribution of the *increase* in tuition during this 9 year period by region.  Be sure to include appropriate labels and titles in your plot.  Discuss how the estimated tuition increase varies by region.

2.  Create a suitable visualization to show the distribution of college type by region.  Be sure to use appropriate labels and titles in your plot.  Describe in a few sentences what you can conclude from your plot.   

3.  What is the minimum expected private college tuition in the Northeast region in the year 2000?  Which region and type of college has the highest median graduation rate?  In the South, what percentage of colleges are private? Be sure to include the code that you used to answer each of these questions.

4.  Choose a numerical and a categorical variable and construct side-by-side box plots depicting the relationship between the two variables (you must choose a pair of variables that is different than the pair of variables you plotted in Question 1).  Describe in a few sentences what you can conclude from your plot.


# Part II: Nobel Laureates

In January 2017, Buzzfeed published an article on why Nobel laureates show immigration is so important for American science.
You can read the article [here](https://www.buzzfeednews.com/article/peteraldhous/immigration-and-science).
In the article they show that while most living Nobel laureates in the sciences are based in the US, many of them were born in other countries.
This is one reason why scientific leaders say that immigration is vital for progress.
In this part of the assignment, we will work with the data from this article to recreate some of their visualizations as well as explore new questions.

The dataset for this assignment can be found as a CSv (comma separated values) file in the `data` folder of your repository.


The variable descriptions are as follows:

-   `id`: ID number
-   `firstname`: First name of laureate
-   `surname`: Surname
-   `year`: Year prize won
-   `category`: Category of prize
-   `affiliation`: Affiliation of laureate
-   `city`: City of laureate in prize year
-   `country`: Country of laureate in prize year
-   `born_date`: Birth date of laureate
-   `died_date`: Death date of laureate
-   `gender`: Gender of laureate
-   `born_city`: City where laureate was born
-   `born_country`: Country where laureate was born
-   `born_country_code`: Code of country where laureate was born
-   `died_city`: City where laureate died
-   `died_country`: Country where laureate died
-   `died_country_code`: Code of country where laureate died
-   `overall_motivation`: Overall motivation for recognition
-   `share`: Number of other winners award is shared with
-   `motivation`: Motivation for recognition

In a few cases the name of the city/country changed after laureate was given (e.g. in 1975 Bosnia and Herzegovina was called the Socialist Federative Republic of Yugoslavia).
In these cases the variables below reflect a different name than their counterparts without the suffix \`\_original\`.

-   `born_country_original`: Original country where laureate was born
-   `born_city_original`: Original city where laureate was born
-   `died_country_original`: Original country where laureate died
-   `died_city_original`: Original city where laureate died
-   `city_original`: Original city where laureate lived at the time of winning the award
-   `country_original`: Original country where laureate lived at the time of winning the award


5.  Read in the dataset `nobel.csv`, which can be found in the `data` folder in your repository. How many observations and how many variables are in the dataset?  **Use inline code to answer this question.**

6.  There are some observations in this dataset that we will exclude from our analysis to match the Buzzfeed results.  Create a new data frame called `nobel_living` that only includes observations (or laureates) such that:

-   `country` is available (that is, the value of `country` is not `NA`)
-   the laureates are people as opposed to organizations (organizations are denoted with `"org"` as their `gender`)
-   the laureates are still alive (that is, their `died_date` is `NA`).  

Confirm that once you have filtered for these characteristics you are left with a data frame with 228 observations, once again using inline code.

7.  According to Buzzfeed:  **Most living Nobel laureates were based in the US when they won their prizes**

You will check to see if that claim is true.

First, create a new variable to identify whether the laureate was in the US when they won their prize using the `mutate()` function.
The following code mutates the `nobel_living` data frame by adding a new variable called `country_us`using an `if` statement.
The first argument in the `if_else()` function we're using to write this if statement is the condition we're testing for.
If `country` is equal to `"USA"`, we set `country_us` to `"USA"`.
If not, we set the `country_us` to `"Other"`.


```{r eval = FALSE}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

Next, we will limit our analysis to only the following categories: Physics, Medicine, Chemistry, and Economics.

```{r eval = FALSE}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

For this exercise work with the `nobel_living_science` data frame created above.  Be sure to remove the `eval = FALSE` option in the two code chunks above.

Create a faceted bar plot visualizing the relationship between the category of prize and whether the laureate was in the US when they won the nobel prize.
    Interpret your visualization, and say a few words about whether the Buzzfeed claim **Most living Nobel laureates were based in the US when they won their prizes** is supported by the data.

    -   Your visualization should be faceted by category.
    -   For each facet you should have two bars, one for winners in the US and one for Other.
    -   Flip the coordinates so the bars are horizontal, not vertical.
    -   Give your visualization an appropriate title and make sure all axis labels are appropriate.


8.  Create a new variable called `born_country_us` in the `nobel_living_science` dataset that has the value `"USA"` if the laureate was born in the US, and `"Other"` otherwise.
How many of the winners are born in the US?

9.  According to Buzzfeed:  **Of those US-based Nobel laureates, many were born in other countries**.

Add another variable to your visualization from Exercise 7 that indicates whether the laureate was born in the US or not.
    Based on your visualization, do the data appear to support Buzzfeed's claim that **Of those US-based Nobel laureates, many were born in other countries**?
    Explain your reasoning in 1-2 sentences.

  
10. **Here's where those immigrant Nobelists were born**

For this exercise work with the `nobel_living_science` data frame created previously.

In a single pipeline, filter for laureates who won their prize in the US, but were born outside of the US, and then create a frequency table (with the `count()` function) for their birth country (`born_country`) and arrange the resulting data frame in descending order of number of observations for each country. Which country is the most common?

## Grading

In addition to accuracy of your answers to questions, your submission will be 
evaluated for

- coding style,
- informatively named code chunks,
- commit after each question (at a minimum, more commits ok),
- informative commit messages,
- document organization, and
- quality of writing and grammar.

## Assignment Submission

Once you have completed the assignment, be sure to create a pdf file from your html file and push it to your HW2 repository.